
# Analysis

## Exploratory Data Analysis (EDA)

The first step of conducting analysis is to understand the data by conducting exploratory data analysis. To conduct the EDA, we obtained descriptive statistics and summaries of all variables. For the quantitative variables, we wrote a function called output_quantitative_stats() to get minimum, maximum, range, median, first and third quartiles, IQR, Mean and Sd of all the quantitative variables including "UGDS_BLACK", "UGDS_HISP", "UGDS_ASIAN", "UGDS_ASIAN", "UGDS_AIAN", "UGDS_NHPI", "UGDS_2MOR","UGDS_NRA", "UGDS_UNKN","UGDS_WHITE", "UGDS", "ADM_RATE", "COSTT4_A", "MD_EARN_WNE_P10", "C100_4", "PCTFLOAN", "CIP_SUM", "MINORATIO", and "STU_APPLIED". 

Similarly, we wrote a function called output_qualitative_stats() to generate a table with both the frequency and the relative frequency of the qualitative variables including "WEST", "MIDWEST", "NORTHEAST", "SOUTH", "MAJOR_CITY", "MINOQ1", "MINOQ2", "MINOQ3",and "MINOQ4". To understand the data better, we also want to generate some plots to visualize the data. We wrote the functions histogram_generator() and boxplot_generator() to generate histograms and boxplots of the quantitative variables and condition_boxplot_generator() to generate conditional boxplots between "STU_APPLIED" and the qualitative variables. To study the association between "STU_APPLIED" and the rest of predictors, we also obtained the correlation matrix of all quantitative variables using function cor(), the scatterplot matrix using function pairs(), the anova between "STU_APPLIED" and all the qualitative variables using function aov().

Then, we divide our data set into 8 separate clusters according to region and its proximity to major cities. In order to develop strategies to improve competitiveness for each cluster, we first tabulate some key statistics for each cluster.

```{r results= 'asis', echo = FALSE}
library(xtable)
library(Matrix)
options(xtable.comment = FALSE,
        xtable.table.placement = "H")
load("../data/data-outputs/eda-outputs/cluster-eda-stats.RData")
print(xtable(cluster_eda_stats, caption = 'Cluster Statistics'), comment = FALSE)
```

The table contains the number of institutions in each cluster, with Northeast region not located in major city cluster having the most instituions. And since we are interested in predicting students applied by variables such as earnings, graduation rate, minority ratio and percentage of students with loans, we look at the mean for this variable in each cluster first. 

ANOVA is designed to test whether there are any statistically significant differences between the means of independent groups. Since our measure of school competitiveness is the number of students applied, we will test whether the means of students applied is different among clusters to gauge whether our clustering criteria makes sense. 

```{r results= 'asis', echo = FALSE}
load("../data/data-outputs/anova.RData")
print(xtable(summary(aov_result), caption = 'ANOVA Test Result'), comment = FALSE)
```

The test result shows that we have a p-value smaller than 0.01, which means that we can reject the null hypothesis that the means are the same across all 8 clusters. 

Then, after the premilinary EDA, we start to run regression and explore the relationship between variables. 

To start with, we run an OLS regression for all variables that we believe have an impact on number of students applied. 

Students applied = Median_Earning + Completion_rate + Percentage_with_Student_Loans + Major_City + Minority_Ratio + West + Midwest + Northeast

The dependent variable is number of students applied in the fall term. It is calculated by dividing people admitted during fall term by the admission rate.

We use 6 main variables as regressors in the regression:

1. Median_Earning: Student's median earning 10 years after graduation

2. Completion_rate: Percentage of students graduated within 4 years. 

3. Percentage_with_Student_Loans: Percentage of students with Student Loans.

4. Major_City: Whether the institution is located near a major city or countryside. This dummy variable equals 1 if it is located in a major city, 0 if countryside. 

5. Minority_Ratio: THe ratio of non-white students to the total population.

6. West, Midwest, East: Region dummy variables. We divided all schools into 4 regions and if an institution belongs to a certain region, the corresponding dummy varible will be 1, and others be 0. We drop one dummy variable Northeast in the multiple linear regression to avoid perfect collinearity. 

